>

### 2019-04-08 paper/spinifex_paper.Rmd
Changed abstract, moved code betweeen steps. consistency. Gammerly and knit.

---
title: > 
  spinifex: manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - School of Information Technology
    - Monash University, Melbourne, VIC 3800, Australia
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - School of Econometrics and Business Statistics
    - Monash University, Melbourne, VIC 3800, Australia
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provide a unique dynamic visualization as the dimensionality of the data increase. Tours are particularly useful for understanding the structure held within multivariate data, and in association with techniques for dimension reduction, supervised, and unsupervised classification. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the contemporary animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed. This particularly useful after identifying a feature of interest.

  A recent paper {@wang_mapping_2018} visualizes the sensitivity of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolve more structural information that is orthogonal to the original viewing plane.
keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
output:
  bookdown::pdf_book:
    #base_format: rticles::rjournal_article #see sandbox for hard code solution.
    fig_caption: yes
    toc: false
    includes:
      in_header: preamble.tex
bibliography: RJrefs.bib
---

@@

<!-- Tours -->
A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity. This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a select variable has in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@cook_grand_1995]. The path of a guided tour is selected via projection pursuit, the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to identify interesting projection features rapidly given the relatively large parameter-space. Once the given projection has been provided, it is time to define the path of rotation.

<!-- Manual tour paths -->
Ideally, the path would be intuitively user-generated from physical movement, be it through mouse or motion capture. Unfortunately, this type of dynamic control has proven difficult to capture for in R. Because of this, manual tours were not implemented in **tourr** [@wickham_tourr_2011]. This research allows for the consumption, but not the generation, of such dynamic input. After the capture of an oblique user motion, the rotation needs to be applied to step 3 (rotation sequence) of the algorithm discussed below. In the section below we stick with a radial rotation where, $\theta$, the angle of in-projection-plane rotation is held constant.

@@

The algorithm primarily operates on projection bases and utilizes the data only when making a display. The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.

@@

<!-- Projection basis -->
The data is defined. A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation. One way to identify a projection containing interesting features is to apply a guided tour [@cook_interactive_2007]. In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space. In this case, the holes index is selected and applied to standardized flea data. The holes index is maximized when the projected observations are furthest from the center. Figure \@ref{fig:step0} shows a locally optimized projection of the data. The left side displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection. The right side shows the data as projected through the basis set described by the reference axes (left). Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).

@@

```{r step0, echo=F, results='hide', out.width='95%', fig.cap = "(ref:step0-cap)"}
#"\\label{fig:step0} and some caption"}

@@

@@

<!-- Zero Vect, manip sp -->
Initialize a zero vector **e** of length *p*. Set the fifth element to on, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.

\begin{align*}
  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~~~ w.r.t. the basis \\
  &= \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

```{r step2, echo=F, out.width='100%', fig.cap = "(ref:step2-cap)"}
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
    theme(panel.border = element_rect(colour = "black", fill = NA))
```

@@

@@

Let:
\begin{description}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi$
\end{description}

@@

<!-- ```{r STEP3CODEEXAMPLE, eval=F, echo=T} -->
<!-- for (phi in seq(seq_start, seq_end, phi_inc_sign)) { -->
<!--   slide <- slide + 1 -->
<!--   tour[,, slide] <- rotate_manip_space(manip_space, theta, phi)[, 1:2] -->
<!-- } -->
<!-- ``` -->

@@

## Data in projection-space {#sec:display}

@@

<!-- grand tours on the same data -->
Using the same data, another study, @cook_dynamical_2018, applies grand tours to the same subspaces. Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermidiate frames. Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than use of PDFSense & TFEP (TensorFlow embedded projections). Before applying manual tours let's discuss the structure of the data.

<!-- Data structure -->
The data has a hierachrical structure with top-level groupings; `DIS`, `VBP`, and `jet`. Each group is a particular class of experiments, each with many experimental datasets which, in turn, have many observations. In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters. This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.

@@

<!-- motivate jet cluster, explain dimensionality  -->
The `jet` cluster is of interest as it contains the largest data sets and is found to be important in @wang_mapping_2018. The jet cluster resides in a smaller dimensionality than the full set of experiments with 4 principal components explaining 95% of the variation in the jet cluster [@cook_dynamical_2018]. The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \@ref(fig:JetClusterGood) and figure \@ref(fig:JetClusterBad) respectively. These components are selected to contrast the difference of information conveyed by touring different variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.

@@

(ref:JetClusterBad-cap) Jet cluster, a radial manual tour of PC4. Colored by experiment type: ATLAS7new in green and ATLAS7old in orange. This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, regardless of the contribution of PC4. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

@@

## TODO: edit application, Grammarly and word checks.
<!-- discussion of findings and which is better -->
Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.

Jet cluster manual tours manipulating each of the principal components can be viewed from the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).

@@

We perform a manual tour on this data, manipulating PC6 as depicted in figure \@ref(fig:DISclusterGood). Looking at several frames we see that DIS HERA data lies mostly on a plane. When PC6 has full contributions, we see the dimuon SIDIS in purple is almost orthogonal to the DIS HERA (green). Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the DIS HERA data. A dynamic version of this manual tour can be found at:
https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
The page may some time to load, as the animation is several megabytes.

(ref:DISclusterGood-cap) DIS cluster, a radial manual tour of PC6. colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange. When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to DIS HERA (green) data. As the projection is rotated, we can also see that DIS HERA (green) practically lies on a plane in this 6D subspace. When the contribution of PC6 is near zero, dimonSIDIS (purple) occupies the same space as the DIS HERA data. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.

@@

(ref:DISclusterBad-cap) DIS cluster, a radial manual tour of PC2. Colored by experiment type: DIS HERA1+2 in green, dimuon SIDIS in purple, and charm SIDIS in orange. The structure of previously described plane of DIS HERA (green) and nearly orthogonal dimuon SIDIS (purple) is present, however, the manipulating PC2 does not give a head-on view of either, a less useful manual tour than that of PC6. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.

@@

## Source code and usage {#sec:usage}

Use the below code as a guide for installation and finding the vignette. The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases. If you prefer to follow along with the example in the algorithm then simplified code is also listed below.

```{r INSTALLCODEEXAMPLE, eval=F, echo=T}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")

## manual tour of std flea from holes-index:
library(spinifex)
f_dat  <- tourr::rescale(flea[,1:6])
f_cat  <- factor(flea$species)
f_path <- save_history(f_dat, guided_tour(holes()))
f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_mvar <- 5
f_lab  <- colnames(f_dat)

# View the basis
view_basis(f_bas, data = f_dat, labels = f_lab)
# View the manip space
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab)
# Play animation as HTML5 from plotly
play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar, 
                 col = f_cat, angle = f_angle)
```

### Acknowledgments

@@

<!-- geometric and graphics display -->
The **tourr** package provides many other geometric displays with the `tourr::display_*()` family. These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**. Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework. However, **animation** builds from base graphs while **spinifex** utilizes **ggplot2** graphics.

@@

>>>>

---
title: > 
  spinifex: manual control of dynamic linear projections of high-dimensional data
author:
  - name: Nicholas Spyrison
    affiliation: Monash University 
    address:
    - School of Information Technology
    - Monash University, Melbourne, VIC 3800, Australia
    email:  Nicholas.Spyrison@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address:
    - School of Econometrics and Business Statistics
    - Monash University, Melbourne, VIC 3800, Australia
    email:  DiCook@monash.edu
abstract: >
  The set of dynamic linear projections of multivariate data collectively known as **tours** provides an important tool for extending the dimensionality of visuals. The R package **tourr** offers a variety of path generators and geometric displays for conducting tours. This paper discusses an extension package, **spinifex**, that adds support for the path generation of manual tours and extends the display of tours to use with the animation packages, **plotly** and **gganimate**. Manual tours are used to explore the sensitivity of structure as the contributions of a manipulation variable are changed.

  A recent paper {@wang_mapping_2018} visualizes the sensitivity of the hadronic experiments to nucleon structure. Sensitivity was characterized in non-linear 3D embeddings of the first 10 principal components. This research applies manual tours to this data showing that manual tours resolve information about cluster structure and identify outliers, information that is orthogonal to the original static viewing plane.
keywords: manual tour, guided tour, grand tour, projection pursuit, high dimensional data, multivariate data, data visualization, statistical graphics, data science
output:
  bookdown::pdf_book:
    #base_format: rticles::rjournal_article #see sandbox for hard code solution.
    fig_caption: yes
    toc: false
    includes:
      in_header: preamble.tex
bibliography: RJrefs.bib
---

@@

<!-- Tours -->
A tour is a multivariate data analysis technique in which a sequence of linear (orthogonal) projections are viewed as an animation while the orientation of the projection basis is rotated across time. Each frame of the sequence corresponds to a small change in the projection for a smooth transition that perseveres continuity. This is a crucial tool extending the dimensionality of visualization which is a large boon as datasets and models exist in increasing cardinality. The package **tourr** [@wickham_tourr_2011] builds a platform for generating tour paths and applying to a wide number of display geometrics in **base** graphics.

<!-- Manual tour -->
While there are numerous methods that generate tour paths, this research focuses on the manual tour. The manual tour was described in @cook_manual_1997 and allows a user to control the projection coefficients of a selected variable in a 2D projection. The manipulation of these coefficients allows the analyst to explore how sensitive the projections structure is to these changes. This makes manual tours particularly useful once a feature of interest has been identified, for example, with the use of a guided tour [@hurley_analyzing_1990]. Rather than exploring the large phase space of the data, a guided tour selects a very specific path via projection pursuit[@diaconis_asymptotics_1984], the optimization of an index function on the projection via a hill climbing algorithm. This allows guided tours to rapidly identify interesting projection features given the relatively large parameter-space. After a projection of interest is identified an analyst can then use the "finer brush" of the manual tour, by controlling the contributions of individual variables.

@@

<!-- operate on bases -->
The algorithm primarily operates on projection bases and utilizes the data only when making a display. The projection space can be viewed at any point in the process by pre-multiplying the data and plotting the first two variables.

@@

<!-- Projection basis -->
The data is defined. A basis set (ideally orienting to an interesting feature) must be provided as an initial orientation. One way to identify a projection containing interesting features is to apply a guided tour [@cook_interactive_2007]. In a guided tour the projection sequence is selected by optimizing an index function via hill-climbing on the projection space. In this case, the holes index is selected and applied to standardized flea data. The holes index is maximized when the projected observations are furthest from the center. Figure \@ref{fig:step0} shows a locally optimized projection of the data. The left panel displays the reference axes of the projection basis, a visual indication of the magnitude and direction each variable contributed to the projection. The right panel shows the data as projected through the basis set described by the reference axes (left). Data points are colored and given shape according to the species (while the guided tour was unsupervised with this information).

@@

```{r step0, echo=F, results='hide', out.width='60%', fig.cap = "(ref:step0-cap)"}
#"\\label{fig:step0} and some caption"}

@@

<!-- step 0 disp code. -->
```{r step0CODEEXAMPLE, eval=F, echo=T}
# devtools::install_github("nspyrison/spinifex") # Development version
install.package("spinifex")
library("spinifex")

f_data  <- tourr::rescale(flea[,1:6])
f_path  <- save_history(f_data, guided_tour(holes()))
f_basis <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2)
f_cat   <- factor(flea$species)

view_basis(basis = f_basis, 
           data = f_data,
           labels = colnames(f_data))
```

@@

<!-- Zero Vect, manip sp -->
Initialize a zero vector **e** of length *p*. Set the fifth element to one, as  `aede2` is the fifth variable in the data, giving the manip var a full contribution in this dimension. Use the Gram-Schmidt process to orthonormalize the zero vector onto the basis yielding the 3D manipulation space, **M**.

\begin{align*}
  \textbf{e} &\leftarrow Orthonormalize_{GS}(\textbf{e}) ~~~ w.r.t.~the~basis \\
  &\leftarrow \textbf{e} - \langle \textbf{e},\textbf{B}_1 \rangle \textbf{B}_1 - \langle \textbf{e}, \textbf{B}_2 \rangle \textbf{B}_2 \\
  \\
  \textbf{M}_{[p,~3]} &= (\textbf{B}_1,\textbf{B}_2,\textbf{e})
\end{align*}

@@

```{r step2, echo=F, out.width='40%', fig.cap = "(ref:step2-cap)"}
view_manip_space(basis = f_bas, manip_var = f_mvar, labels = colnames(f_dat)) +
    theme(panel.border = element_rect(colour = "black", fill = NA))
```

@@

<!-- step 2 disp code. -->
```{r step2CODEEXAMPLE, eval=F, echo=T}
view_manip_space(basis = f_basis, 
                 manip_var = 5, 
                 labels = colnames(f_data))
```

@@

Let:
\begin{itemize}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi$
\end{itemize}

@@

@@

<!-- step 3 disp code. -->
```{r step3CODEEXAMPLE, eval=F, echo=T}
angle_speed <- .26

play_manual_tour(data = f_data,
                 basis = f_basis, 
                 manip_var = 5, 
                 angle = angle_speed,
                 col = f_cat,
                 pch = f_cat)
```

### Data in projection-space {#sec:display}

@@

<!-- grand tours on the same data -->
Using the same data, another study, @cook_dynamical_2018, applies grand tours [@asimov_grand_1985] to the same subspaces. Grand tours are dynamic subspace projections of high dimensional where frames and selected at random and linked with geodesically interpolation of the intermediate frames. Grand tours are able to better resolve the distribution shape of clusters, intra-cluster detail, and better outlier detection than the use of PDFSense & TFEP (TensorFlow embedded projections). Before applying manual tours let's discuss the structure of the data.

<!-- Data structure -->
The data has a hierarchical structure with top-level clusters; `DIS`, `VBP`, and `jet`. Each cluster is a particular class of experiments, each with many experimental datasets which, in turn, have many observations. In the consideration of data density, we conduct manual tours on subsets of the `DIS` and `jet` clusters. This explores the sensitivity of the structure to each of the variables in turn and we present the subjectively best and worst variable to manipulate for identifying dimensionality of the clusters and describing the span of the clusters.

@@

<!-- jet cluster, explain dimensionality  -->
The `jet` cluster resides in a smaller dimensionality than the full set of experiments with four principal components explaining 95% of the variation in the cluster [@cook_dynamical_2018]. The data within this 4D embedding is subset down to `ATLAS7old` and `ATLAS7new` to focus in on two groups with a reasonable number of observations that occupy different parts of the subspace. Radial manual tours controlling contributions from PC4 and PC3 are shown in figure \@ref(fig:JetClusterGood) and figure \@ref(fig:JetClusterBad) respectively. These components are selected to contrast the difference of information conveyed by touring different variables. Links to dynamic HTML5 animations controlling each of the four variables are also provided.

@@

(ref:JetClusterBad-cap) `Jet` cluster, a radial manual tour of PC4. Colored by experiment type: ATLAS7new in green and ATLAS7old in orange. This manual tour contains less interesting information ATLAS7new (green) has points that are right and left of ATLAS7old, while most points occupy the same projection space, independent of the contribution of PC4. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/.

@@

## TODO: edit DIS cluster, Grammarly and word checks.
<!-- discussion of findings and which is more insightful -->
Manipulating PC3, where varying the angle of rotation brings interesting features into and out of the center mass of the data, is more interesting than the manipulation of PC4, where the features are mostly independent of the contribution of PC4.

Radial manual tours manipulating each of the principal components in the `jet` cluster can be viewed by following the links: [PC1](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc1/), [PC2](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc2/), [PC3](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc3/), and [PC4](https://nspyrison.netlify.com/thesis/jetcluster_manualtour_pc4/).

@@

We perform a manual tour on this data, manipulating PC6 as depicted in figure \@ref(fig:DISclusterGood). Looking at several frames we see that `DIS HERA` (green) data lie mostly on a plane. When PC6 has full contributions, `dimuon SIDIS` (purple) is almost orthogonal to the `DIS HERA` (green). Yet the contribution of PC6 has zeroed the dimuon SIDIS data occupy the same space as the `DIS HERA` data. A dynamic version of this manual tour can be found at:
https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.
The page may some time to load, as the animation is several megabytes.

(ref:DISclusterGood-cap) `DIS` cluster, a radial manual tour of PC6. colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange. When the contribution PC 6 is large we see that dimuon SIDIS (purple) data are nearly orthogonal to `DIS HERA` (green) data. As the projection is rotated, we can also see that `DIS HERA` (green) practically lies on a plane in this 6D subspace. When the contribution of PC6 is near zero, `dimonSIDIS` (purple) occupies the same space as the `DIS HERA` data. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc6/.

@@

(ref:DISclusterBad-cap) `DIS` cluster, a radial manual tour of PC2. Colored by experiment type: `DIS HERA1+2` in green, `dimuon SIDIS` in purple, and `charm SIDIS` in orange. The plane of `DIS HERA` (green) and nearly orthogonal `dimuon SIDIS` (purple) is present, however, manipulating PC2 does not give a head-on view of either, offers less information about these features than manipulating PC6. A dynamic version can be viewed at https://nspyrison.netlify.com/thesis/discluster_manualtour_pc2/.

@@

<!-- ## Source code and usage {#sec:usage} -->

<!-- Use the below code as a guide for installation and finding the vignette. The vignette offers a less technical discussion opting to focus on code usage and goes through a couple more use cases. If you prefer to follow along with the example in the algorithm then simplified code is also listed below. -->

<!-- ```{r INSTALLCODEEXAMPLE, eval=F, echo=T} -->
<!-- # devtools::install_github("nspyrison/spinifex") # Development version -->
<!-- install.package("spinifex") -->
<!-- library("spinifex") -->

<!-- # Also see vignette: -->
<!-- vignette("spinifex") # vignette ‘spinifex’ not found -->

<!-- ## manual tour of std flea from holes-index: -->
<!-- f_dat  <- tourr::rescale(flea[,1:6]) -->
<!-- f_cat  <- factor(flea$species) -->
<!-- f_path <- save_history(f_dat, guided_tour(holes())) -->
<!-- f_bas  <- matrix(f_path[,, max(dim(f_path)[3])], ncol=2) -->
<!-- f_mvar <- 5 -->
<!-- f_lab  <- colnames(f_dat) -->

<!-- # View the basis -->
<!-- view_basis(f_bas, data = f_dat, labels = f_lab) -->
<!-- # View the manip space -->
<!-- view_manip_space(basis = f_bas, manip_var = f_mvar, labels = f_lab) -->
<!-- # Play animation as HTML5 from plotly -->
<!-- play_manual_tour(data = f_dat, basis = f_bas, manip_var = f_mvar,  -->
<!--                  col = f_cat, angle = f_angle) -->
<!-- ``` -->

@@

<!-- geometric and graphics display -->
The **tourr** package provides many other geometric displays with the `tourr::display_*()` family. These geometric options could be integrated into the **ggplot2** framework for display on **plotly** and **gganimate**. Additionally, the **animation** package @xie_animation:_2018 could be implemented for another graphics framework. However, **animation** builds from **base** graphs while **spinifex** utilizes **ggplot2** graphics.

@@

## Acknowledgments
